---
title: "Gene Expression Analysis"
author: "wlw323"
date: "`r Sys.Date()`"
output: 
  workflowr::wflow_html:
     toc: false
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Gene Expression Analysis

This document contains the analysis of gene expression data using Cuffdiff results.

```{r read-data, results='hide'}
file_path <- "c:/Users/wlw/Desktop/sx/gene_exp.diff"
cuffdiff_result <- read.table(file = file_path, header = TRUE)
```

### Calculating Log2 Fold Change and Adjusting for Zero Values

```{r calculate-log2-foldchange, results='hide'}
wt_FPKM <- cuffdiff_result$value_1
tre_FPKM <- cuffdiff_result$value_2
log2_foldchange <- log2(tre_FPKM / wt_FPKM)
log2_foldchange[tre_FPKM == 0] <- 0
log2_foldchange[wt_FPKM == 0] <- 0
```

### Calculating Negative Log10 P-Value

```{r calculate-log10-p-value, results='hide'}
log10_p_value <- log10(cuffdiff_result$p_value) * -1
```

### 1st Edition: Displaying the Original Plot

```{r plot-original, fig.width=8, fig.height=6}
plot(x = log2_foldchange, y = log10_p_value)
```

### 2nd Edition: Adjusting Plot Limits and Filtering Data Points

```{r plot-filtered, fig.width=8, fig.height=6}
log10_p_value.filter <- log10_p_value[log10_p_value >= 0.001]
log2_foldchange.filter <- log2_foldchange[log10_p_value >= 0.001]
plot(x = log2_foldchange.filter, y = log10_p_value.filter, 
     xlim = c(-5, 5), ylim = c(0.225, 4.5))
```

### 3rd Edition: Coloring the Plot

```{r plot-colored, fig.width=8, fig.height=6}
# Selecting significant points
selec_sign_vector <- (cuffdiff_result$value_1 > 0) & (cuffdiff_result$value_2 > 0) &
                     (cuffdiff_result$value_1 >= 1 | cuffdiff_result$value_2 >= 1) &
                     (cuffdiff_result$p_value <= 0.05) & (abs(log2_foldchange) >= 1)
selec_sign_vector.filter <- selec_sign_vector[log10_p_value >= 0.001]
col_vector <- rep(rgb(0, 0, 1, 0.1), length(log10_p_value.filter))
col_vector[selec_sign_vector.filter] <- rgb(1, 0, 0)

# Filtering for plot
log10_p_value.filter <- log10_p_value[log10_p_value >= 0.001]
log2_foldchange.filter <- log2_foldchange[log10_p_value >= 0.001]
plot(x = log2_foldchange.filter, y = log10_p_value.filter,
     xlim = c(-4, 4), ylim = c(0.001, 4.5), col = col_vector, pch = 16)

# Adding horizontal line at -log10(0.05)
abline(h = -1 * log10(0.05), lwd = 3, lty = 3, col = "#4C5B61")
```

This document provides a step-by-step analysis of gene expression data, including data filtering, calculation of log2 fold change, and visualization of the results with significant points highlighted.
```


